 Casting multiple shadows: high-dimensional interactive data visualisation with tours and embeddings

### 2020-07-18 paper.Rmd
update code and text

The next section steps through three use cases of our approach using simulated
and real data.  The videos presented show the visual 
appearance of the `liminal` interface, and how we can interact with the tour
via the controls previously described.
In the first case study, we look at three different examples
where t-SNE is applied to three simulations where cluster structure and geometry
is known with varying success.

In the second case study, we explore a well-clustered single cell
transcriptomics data of mice retina...

@@

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the `scRNAseq` Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the the `scater`
by removing cells with high proportion of mitochondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using `scran` 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering to generate clusters [@Blondel2008-bx]. 
To check and verify the clustering we construct a liminal view. Because most off the variability in
expression is explained by the first 5 PCs, we tour those. The t-SNE view
is computed from all PCs.

>>>>

The next section steps through three use cases of our approach using simulated
and real data.  The videos presented show the visual 
appearance of the `liminal` interface, and how we can interact with the tour
via the controls previously described.
In the first case study, we look at three different examples
where t-SNE is applied to three simulations where cluster structure and geometry
of the underlying data is known.

In the second case study, we look at well-clustered single cell
transcriptomics data of mice retina...

@@

First, we downloaded the raw mouse retinal single cell RNA-seq data from
@Macosko2015-ot using the `scRNAseq` Bioconductor package [@scRNAseq-d]. We
have followed a standard workflow for pre-processing and normalizing this data 
(described by @Amezquita2020-at): we performed QC using the the `scater`
by removing cells with high proportion of mitochondrial gene expression and low
numbers of genes detected, we log-transformed and normalised the expression 
values and  finally selected highly variable genes (HVGs) using `scran` 
[@McCarthy2017; @Lun2016]. The top ten percent of HVGs were used to subset
the normalised expression matrix and compute PCA using the first 25 components.
Using the PCs we built a shared nearest neighbours graph (with $k = 10$) and 
used Louvain clustering to generate clusters [@Blondel2008-bx]. 
To check and verify the clustering we construct a liminal view. 
We tour the first five PCs (approximately 20% of the variance in expression),
along side the tSNE view which was computed from all 25 PCs. Due to latency
of the interface we do a ten percent weighted sample of rows based on
cluster membership.
